Comment: Reasoning about Interference Between Units

### 2016-03-26 paper/introduction.tex
Because one of the reviewers pointed out typos, this commit is the result of fixing some of them.

Notice that the test statistic choice matters in this process: the engine of
statistical inference involves summarizing information against the
hypothesized claim, yet different ways to summarize information might be more
or less sensitive to substantively meaningful differences. The statistical
power of a simple test of the sharp null hypothesis of no effects will vary as
a function of the design of the study (total number of observations, proportion treated, blocking structure,
etc), characteristics of the outcome (continuous, binary, skewed, extreme
points, etc), and the way that a test statistic summarizes the outcome (does
it compare means, standard deviations, medians, qqplots, etc).\footnote{Some
  use the term `effective sample size' --- which we first saw in
  \citet{kish65} ---
  to highlight the fact that statistical power depends on more than the number
  of rows in a given rectangular dataset. } In general,
test statistics should be powerful against relevant alternatives. \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
advice about the large sample performance of certain classes of test statistics
and BFP repeat his general advice: ``Select a test statistic [$\mathcal{T}$] that will be
small when the treated and control distributions in the adjusted data \ldots
are similar, and large when the distributions diverge.'' \cite[Proposition 4
and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
with this property (``effect increasing'' test statistics), produce an
unbiased test of the hypothesis of no effects or positive effects when the
positive effects involve one parameter. Such results mean that a test of
the sharp null of no effects
based on a known randomization using \emph{any} effect increasing
test-statistic to be valid test (in that the test should produce $p$-values
less than $\alpha$ no more than 100$\alpha \%$ of the time when the null is true) even though different
test statistics may imply different power against false hypotheses.  Yet, when the models are complex and
may involve increasing effects in the direction of one parameter and
non-linear effects in the direction of another parameter, BFP showed that
sometimes a KS-test will have (1) no power to address the model at all such
that all hypothesized parameters would receive the same implausibility
assessment; (2) or might reject all such hypothesized parameters. Thus,
although in theory one may assess sharp multiparameter hypotheses,
 in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
characteristics of tests as a piece of their workflow --- because  the theory
justifying simple one-dimensional effect increasing test statistics clearly
did not cover multi-parameter situations like those easily arising from social
network experiments.

@@

Fisher/Rosenbaum style randomization inference tends to use test statistics that
compare two distributions. Simple models imply that the distribution of the
outcome in the control remains fixed. For example, $\widetilde
y_{i,Z_i=0}=Y_i-Z_i \tau$ only changes the distribution of outcomes for units
in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=1}|Z_i=1$
to the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$ makes intuitive sense in this
case, and, if $Y_i$ is Normal or at least unimodal without major outliers,
then this test using means as test statistics might have optimal power. The complex model used as an example
by BFP involved adjustments to both control and treated outcomes --- some
hypothesized parameters would cause shifts in variance, others in location.
So, BFP proposed to use the KS-test statistic to assess the relationship
between $\widetilde y_{i,Z_i=0,Z_{-i}=0}$ and $Z_i$ (where $Z_{-i}=0$ means
``when all units other than $i$ are also not treated".)

Yet, one can also think about the process of hypothesis testing as a process
of assessing model fit, and there are usually better ways to evaluate the fit
of a model than comparing two marginal distributions: For example, the KS test
  uses the maximum difference in the empirical cumulative distributions of
  each treatment group calculated without regard for the relationship between
the treated and control distributions, thereby ignoring information that could
increase the precision of the test. The simplest version of the SSR test
statistic merely sums the difference between the mean of the outcome implied
by the hypothesis and individual outcomes thereby including the correlation
between treated and control outcomes as a part of the distribution of the
statistic:

@@

We compared two models of effects --- a constant additive effects model in
which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant multiplicative effects
model in which $y_{i,Z_i=1}=y_{i,Z_i=0} \cdot \tau$. We set the truth to be
the sharp null of no effects such that the true $\tau=0$ for the additive
model and the true $\tau=1$ for the multiplicative model. To further explain
the process of hypothesis testing and evaluation of the test statistics, we
display the results from a power analysis of one alternative hypothesis in
figure~\ref{fig:boxplot}. A given model and hypothesized parameter implies a
distribution for observed outcomes in the treatment group and control groups.
Here, since we have no interference, the hypotheses only have different
implications for the control group distribution. In this case, we chose a two
parameters for the additive and multiplicative models which produced very
similar implications (as show by the similarity in the ``Add. Model'' and
``Mult.  Model'' boxplots in both panels of the figure). We see that when we
tested the hypothesis of $\tau_0=0.34$ from the additive model using the KS
test statistic on Normal data, that we produced $p$-values lower than
$\alpha=0.05$ about 55\% of the time across 1000 simulations. That is, the KS
test has a power of about 0.55 to reject a false hypothesis with this Normal
outcome and this constant additive effects model. The analogous power for the
SSR test statistic was 0.73. For these particular parameters, we see the SSR
performing better with Normal outcomes but not being very effected by the
model of effects (which makes some sense here because we are choosing
parameters at which both models imply more or less the same patterns in
outcomes). On non-Normal outcomes (shown in the "Jittered Geometric Outcomes"
panel), the SSR test statistic had less power. Again, the particular model of
effect did not matter in this case because we chose the alternative hypothesis
($\tau_0$ on the plots) to represent the case where both models implied
similar outcome distributions.

@@

When we assessed power across many alternative hypotheses, our intuition was
that the SSR would have more power than the KS test when the outcome was
Normal and when the observational implication of the model of effects would
shift means (i.e.\ the additive model). We used direct simulation of the
randomization distribution to generate $p$-values and repeated that process
1000 times to gauge the proportion of rejections of a range of false
hypotheses  (i.e.\ the power of the tests at many different values of
$\tau_0$). The results, not shown here but part of the reproduction archive
for this paper, bear out this intuition: the SSR has slightly more power than
KS for Normal outcomes in both the additive and multiplicative effects
conditions. SSR has slightly less power than KS when the outcome is skewed for
both models. In general, the SSR ought to be most powerful when the effect of
the experiment involves a shift in the location of the distributions of the
treated versus the controls. And it should be less powerful when the effect of
the experiment is mostly to leave the center of the distribution alone, but,
instead to stretch or compress the tails, or to concentrate the experimental
effect on one quantile.

@@

In the case where we know the fixed adjacency matrix of a social network,
$\bS$, and where we imagine that network attributes (like degree) of a node
play a role in the mechanism by which treatment propagates, the idea of
assessing model fit rather than closeness of distributions leads naturally to
the sum-of-squared-residuals (SSR) from a least squares regression of $
\widetilde y_{i,Z_i=0}$ on $Z_{i}$ and $\bz^{T} \bS$ (i.e.\ the number of
directly connected nodes assigned treatment) as well as the $\mathbf{1}^{T}
\bS$ (i.e.\ the degree of the node). If we collected $Z_{i}$,$\bz^{T} \bS$,
and $\mathbf{1}^{T} \bS$ into a matrix $\bX$, and fit the $\widetilde
y_{i,Z_i=0}$ as a linear function of $\bX$ with coefficients $\bbeta$ when we
could define the test statistic as:

@@

Briefly, this model posits that treatment effects can depend on either direct
assignment to treatment ($\bz$) governed by $\beta$ or spillover as an
increasing (but flattening) function of the number of directly connected
treated neighbors ($\bz^{T} \bS$) and is governed by $\tau$. So, we have a
model with two parameters. The network used by BFP involves 256 nodes
connected in an undirected, random graph with node degree ranging from 0 to 10
(mean degree 4, 95\% of nodes with degree between 1 and 8, five nodes with
degree 0 [i.e.\ unconnected]).  Treatment is assigned to 50\% of the nodes 
completely at random in the BFP example.

@@

We cannot say here whether the SSR+Design test will provide the best power
against relevant alternatives for all possible models of treatment effect
propagation, network topologies and designs: this paper uses two models of
effects each of which was applied to a different network topology and
experimental design.\footnote{The BFP paper itself engages with some questions about
the performance of this approach when the theoretical model is very different
from the process generating the data, and we encourage readers to see that
discussion in their \S~5.2.} However, we hope that the results from the 
examples presented here improves the application of the BFP approach and raises new
questions for research.  BFP are correct in the assertion that, regardless of
the choice of test statistic selection, a set of implausible hypotheses is
identified by the procedure. But we should not be led to believe that, for any
given test statistic, that some hypotheses are universally more plausible than others.
Such inferences --- comparing hypotheses --- may depend on the test statistic
used, and not necessarily reflect the plausibility of the model at hand. That
is, the results of any hypothesis test (or confidence interval creation) tell
us \emph{both} about the test statistic \emph{and} about the causal model under scrutiny.

In the example shown in Figure~\ref{fig:twoD}, the SSR+Design test statistic had much better power than
any other test statistic. But SSR from an ordinary least squares regression is
not always appropriate: for example, when the probability of exposure to
spillover is heterogeneous across individuals in a way not well captured by the
$\bz^T \bS$ term or some other analogous term, we may wish to apply inverse
probability weights so as to ensure representative samples of potential
outcomes. This suggests a conjecture: that the $SSR$ from an \emph{inverse-probability-weighted} least squares regression is more generally a
sensible test statistic for models that include
interference.\footnote{\citet{aronowsamii2012interfere} use such weights for
  unbiased estimation of network-treatment-exposure probability weighted
  average treatment effects.}  Additionally, when nonlinear deviations from
model predictions are of concern, a weighted variant of the Brownian distance
covariance \citep{szekely2009brownian} or other $E$-statistic may be more
sensible than the sum of squared residuals.

>>>>

Notice that the test statistic choice matters in this process: the engine of
statistical inference involves summarizing information against the
hypothesized claim, yet different ways to summarize information might be more
or less sensitive to substantively meaningful differences. The statistical
power of a simple test of the sharp null hypothesis of no effects will vary as
a function of the design of the study (e.g. total number of observations, proportion treated, blocking structure), characteristics of the outcome (e.g. continuous, binary, skewed, extreme
points), and the way that a test statistic summarizes the outcome (does
it compare means, standard deviations, medians, qqplots, or something else?).\footnote{Some
  use the term `effective sample size' --- which we first saw in
  \citet{kish65} ---
  to highlight the fact that statistical power depends on more than the number
  of rows in a given rectangular dataset. } In general,
test statistics should be powerful against relevant alternatives. \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
advice about the large sample performance of certain classes of test statistics
and BFP repeat his general advice: ``Select a test statistic [$\mathcal{T}$] that will be
small when the treated and control distributions in the adjusted data \ldots
are similar, and large when the distributions diverge.'' \citet[Proposition 4
and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
with this property (``effect increasing'' test statistics), produce an
unbiased test of the hypothesis of no effects or positive effects when the
positive effects involve one parameter. Such results mean that a test of
the sharp null of no effects
based on a known randomization using \emph{any} effect increasing
test-statistic to be valid test (in that the test should produce $p$-values
less than $\alpha$ no more than 100$\alpha \%$ of the time when the null is true) even though different
test statistics may imply different power against false hypotheses.  Yet, when the models are complex and
may involve increasing effects in the direction of one parameter and
non-linear effects in the direction of another parameter, BFP showed that
sometimes a KS-test will have (1) no power to address the model at all such
that all hypothesized parameters would receive the same high $p$-value; (2) or might describe all such hypothesized parameters as
implausible. Thus,
although in theory one may assess sharp multiparameter hypotheses,
 in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
characteristics of tests as a piece of their workflow --- because  the theory
justifying simple one-dimensional effect increasing test statistics clearly
did not cover multi-parameter situations like those easily arising from social
network experiments.

@@

Fisher/Rosenbaum style randomization inference tends to use test statistics that
compare two distributions. Simple models imply that the distribution of the
outcome in the control remains fixed. For example, the implication of the
constant, additive effects model, $\widetilde
y_{i,Z_i=0}=Y_i-Z_i \tau$, only changes the distribution of outcomes for units
in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=1}|Z_i=1$
to the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$ makes intuitive sense in this
case, and, if $Y_i$ is Normal or at least unimodal without major outliers,
then this test using means as test statistics might have optimal power. The complex model used as an example
by BFP involved adjustments to both control and treated outcomes --- some
hypothesized parameters would cause shifts in variance, others in location.
So, BFP proposed to use the KS-test statistic to assess the relationship
between $\widetilde y_{i,Z_i=0,Z_{-i}=0}$ and $Z_i$ (where $Z_{-i}=0$ means
``when all units other than $i$ are also not treated".)

Although thinking about test statistics as comparing distributions is natural
and simple, one can also think about the process of hypothesis testing as a process
of assessing model fit, and there are usually better ways to evaluate the fit
of a model than comparing two marginal distributions: For example, the KS test
  uses the maximum difference in the empirical cumulative distributions of
  each treatment group calculated without regard for the relationship between
the treated and control distributions, thereby ignoring information about the
joint distribution that could
increase the precision of the test. The simplest version of the SSR test
statistic merely sums the difference between the mean of the outcome implied
by the hypothesis and individual outcomes thereby including the correlation
between treated and control outcomes as a part of the distribution of the
statistic:

@@

We compared two models of effects --- a constant additive effects model in
which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant multiplicative effects
model in which $y_{i,Z_i=1}=y_{i,Z_i=0} \cdot \tau$. We set the truth to be
the sharp null of no effects such that the true $\tau=0$ for the additive
model and the true $\tau=1$ for the multiplicative model. To further explain
the process of hypothesis testing and evaluation of the test statistics, we
display the results from a power analysis of one alternative hypothesis in
figure~\ref{fig:boxplot}. A given model and hypothesized parameter implies a
distribution for observed outcomes in the treatment group and control groups.
Here, since we have no interference, the hypotheses only have different
implications for the control group distribution. In this case, we chose a two
parameters for the additive and multiplicative models which produced very
similar implications (as show by the similarity in the ``Add. Model'' and
``Mult.  Model'' boxplots in both panels of the figure). We see that when we
tested the hypothesis of $\tau_0=0.34$ from the additive model using the KS
test statistic on Normal data, that we produced $p$-values lower than
$\alpha=0.05$ about 54\% of the time across 1000 simulations. That is, the KS
test has a power of about 0.54 to reject a false hypothesis with this Normal
outcome and this constant additive effects model. The analogous power for the
SSR test statistic was 0.70. For these particular parameters, we see the
performance of the SSR test statistic as better than the KS test statistic
with Normal outcomes, but not very effected by the
model of effects (which makes some sense here because we are choosing
parameters at which both models imply more or less the same patterns in
outcomes). On non-Normal outcomes (shown in the ``Jittered Geometric Outcomes"
panel), the SSR test statistic had less power. Again, the particular model of
effect did not matter in this case because we chose the alternative hypothesis
($\tau_0$ on the plots) to represent the case where both models implied
similar outcome distributions.

@@

When we assessed power across many alternative hypotheses in results not shown
here but part of the reproduction archive
for this paper, our intuition was
that the SSR would have more power than the KS test when the outcome was
Normal and when the observational implication of the model of effects would
shift means (i.e.\ the additive model). We used direct simulation of the
randomization distribution to generate $p$-values and repeated that process
1000 times to gauge the proportion of rejections of a range of false
hypotheses  (i.e.\ the power of the tests at many different values of
$\tau_0$). The results bear out this intuition: the SSR has slightly more power than
KS for Normal outcomes in both the additive and multiplicative effects
conditions. SSR has slightly less power than KS when the outcome is skewed for
both models. In general, the SSR ought to be most powerful when the effect of
the experiment involves a shift in the location of the distributions of the
treated versus the controls. And it should be less powerful when the effect of
the experiment is mostly to leave the center of the distribution alone, but,
instead to stretch or compress the tails, or to concentrate the experimental
effect on one quantile.

@@

In the case where we know the fixed binary adjacency matrix of a social
network, $\bS$, where a given entry in the $n \times n$ matrix is 1 if two
units are adjacent and 0 if two units are not adjacent, and where we imagine
that network attributes (like degree) of a node play a role in the mechanism
by which treatment propagates, the idea of assessing model fit rather than
closeness of distributions leads naturally to the sum-of-squared-residuals
(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$
and $\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned
treatment) as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node).
If we collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix
$\bX$, and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
coefficients $\bbeta$ when we could define the test statistic as:

@@

Briefly, this model posits that treatment effects can depend on either direct
assignment to treatment ($\bz$) governed by $\beta$ or spillover as an
increasing (but flattening) function of the number of directly connected
treated neighbors ($\bz^{T} \bS$) and is governed by $\tau$. So, we have a
model with two parameters. The network used by BFP involves 256 nodes
connected in an undirected, random graph with node degree ranging from 0 to 10
(mean degree 4, 95\% of nodes with degree between 1 and 8, five unconnected nodes with
degree 0).  Treatment is assigned to 50\% of the nodes
completely at random in the BFP example.

@@

We cannot say here whether the SSR+Design test will provide the best power
against relevant alternatives for all possible models of treatment effect
propagation, network topologies and designs: this paper uses two models of
effects each of which was applied to a different network topology and
experimental design.\footnote{The BFP paper itself engages with some questions about
the performance of this approach when the theoretical model is very different
from the process generating the data, and we encourage readers to see that
discussion in their \S~5.2.} However, we hope that the results from the
examples presented here improve the application of the BFP approach and raise new
questions for research.  BFP are correct in the assertion that, regardless of
the choice of test statistic selection, a set of implausible hypotheses is
identified by the procedure. But we should not be led to believe that, for any
given test statistic, that some hypotheses are universally more plausible than others.
Such inferences --- comparing hypotheses --- may depend on the test statistic
used, and not necessarily reflect the plausibility of the model at hand. That
is, the results of any hypothesis test (or confidence interval) tell
us \emph{both} about the test statistic \emph{and} about the causal model under scrutiny.

In the example shown in Figure~\ref{fig:twoD}, the SSR+Design test statistic had much better power than
any other test statistic. But SSR from an ordinary least squares regression is
not always appropriate: for example, when the probability of exposure to
spillover is heterogeneous across individuals in a way not well captured by the
$\bz^T \bS$ term or some other analogous term, we may wish to apply inverse
probability weights so as to ensure representative samples of potential
outcomes. This suggests a conjecture: that the $SSR$ from an
\emph{inverse-probability-weighted} least squares regression is a more generally
sensible test statistic for models that include
interference.\footnote{\citet{aronowsamii2012interfere} use such weights for
  unbiased estimation of network-treatment-exposure probability weighted
  average treatment effects.}  Additionally, when nonlinear deviations from
model predictions are of concern, a weighted variant of the Brownian distance
covariance \citep{szekely2009brownian} or other $E$-statistic may be more
sensible than the sum of squared residuals.
